\documentclass[11pt,a4paper]{report}

\usepackage[english]{babel} 
\usepackage[utf8]{inputenc} % Unicode
\usepackage{graphicx} 
\usepackage{url} 
\usepackage{amsmath}
\usepackage{textcomp} 
\usepackage{lipsum}
\usepackage{multirow} 
\usepackage[pdftex,hidelinks]{hyperref}
\usepackage{listings}
\usepackage{natbib}
\usepackage{float}% If comment this, figure moves to Page 2
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{acronym}


\def\R{{\textsl{R }}}
\def\Shiny{\textsl{Shiny }}
\def\sar{\textsl{SARS-CoV-2 }}



\title{Development of a \Shiny application to visualize SARS-CoV-2 vaccination data of Portugal}

\author{
  Bruna Araújo\\
  \texttt{A84408}
  \and
  Matias Capitão\\
  \texttt{A82726} \\ 
  \\ Mentoring:\\
  Cecília Castro\\
  \and
  Rafael Antunes\\
  \texttt{A77457}
}
\date{May or June 2021}



\begin{document}

\maketitle

\begin{abstract}
% Apagar daqui
    ************* Ao fim retornamos aqui.\\ Entretanto encontrei isto para nos ajudar a fazer o Abstract *************\\ \\
%até aqui
    Write the abstract at the very end, when you’ve completed the rest of the text. There are four things you need to include:\\ \\

    Your research problem and objectives\\
    Your methods\\
    Your key results or arguments\\
    Your conclusion

\end{abstract}

\tableofcontents
\newpage

\listoffigures
\newpage

\listoftables
\newpage

{\huge\textbf{Acronym List}}\\
\input{acro_list}


\chapter{Introduction}
\section{\R and \Shiny}

\input{RandShiny}


\section{SARS-CoV-2}
\subsection{Brief description}

\input{Sars-CoV-2}

\subsection{Vaccination}
\input{Vaccination}


\section{Portugal}

\subsection{Brief description}
\input{Portugal}

\subsection{Regions}
\input{RegionsPT}





\section{SARS-CoV-2 in Portugal}

\subsection{Brief description}
\input{CovidPT}

\subsection{Vaccination}
\input{VaccinationPT}





\chapter{Application}

\section{Research}
\input{Research}


\section{Development}
\input{Development}

\subsection{Reactivity in shiny}
\input{ReactivityInShiny}

\subsection{Render functions}

The main function for generating documents in R Markdown, from the rmarkdown package, is a render ().
The render () function is a wrapper that internally calls knitr :: knit () and later converts the document to .html using Pandoc, which is a document and open source converter.Uma função render que usamos na aplicação foi a renderPlot().

To display the graphics in our application we use the renderPlot () tool. RenderPlot is a reactive function that can take input data from the ui.R script and feed it into the server.R script. It then actively updates the information within its role.



\subsection{Data visualization}
Data visualization through maps, tables or graphs allows contextualizing and understanding information in a natural and efficient way.


R has libraries that allow you to play high-quality graphics. In this work, the ggplot2 package was essentially used, whose methods are based on its own grammar, and where the graphical representations are obtained through the (organized) superposition of several levels that are stacked on top of each other, allowing great flexibility in the construction of the graphs (Hadley Wickham, 2005).


Levels, or layers, consist of data (data.frame format), aesthetic attributes (data mapping definition), geometric objects (visual elements), facets (facets()) functions that allow the execution of multiple graphics related to segments of data defined by factors), statistical transformations, coordinates (space where data is represented) and theme (theme(), functions allow changing individual components of a theme).





\section{Implementation}
\subsection{Challenges}

During the development of the App there were a few bumpy moments whether due to the fact that we thought that we were handling the data poorly or when there was a small error that was preventing us from seeing the final result.
Data related, one of the major problems that we encountered was that we could not find a reliable source of population. Every single one we came across did not match up with most of the already existing info on this topic. But after a few weeks we finally came across a document from ECDC, not very well identified, but definitely reliable.
Code related, there were several situations, that we solved quite rapidly, however there was one that stood out the most, namely, when the reports were incomplete. When this happened we always tried to replicate the data with empty values manually, however this solution would not solve our problem completely. The solution we came across was using the merge function so that the weeks that came empty were automatically fulfilled with 'NA' and after that we converted those NA's to 0's. This way we never had a similar problem anymore.

\subsection{User Interface}
\input{userInterface}

\subsection{Portugal}
\input{Rafa}

\input{Matias}

\section{Publication}
Amanhã preencho esta cena depois de publicar

\chapter{Conclusion}

This project made us realize the versatility of the Shiny package from RStudio. We were baffled by how easy it was to build a very useful app with such little tools. Even though we added a bit of HTML and CSS ourselves it was clear that someone with very little knowledge of front-end development and some R language knowledge could very well develop an interesting app.

\bibliographystyle{plain}
\begin{thebibliography}{10}


\bibitem{R-project website} \textit{\href{https://www.r-project.org/about.html}{https://www.r-project.org/about.html}} , \textbf{About/R-Project website}

\bibitem{shiny}\textit{\href{https://shiny.rstudio.com/l}{https://shiny.rstudio.com}} , \textbf{Shiny website
}
\bibitem{SARS-CoV-2}\textit{\href{https://apps.who.int/iris/bitstream/handle/10665/332197/WHO-2019-nCoV-FAQ-Virus_origin-2020.1-eng.pdf}{https://apps.who.int/.../origin-2020.1-eng.pdf}} ,\textbf{ Origin of SARS-CoV-2 - \ac{WHO}}

\bibitem{EU}\textit{\href{https://ec.europa.eu/info/live-work-travel-eu/coronavirus-response/public-health/eu-vaccines-strategy_pt}{https://ec.europa.eu/info/live-work-travel-eu/coronavirus-response/public-health/eu-vaccines-strategy_pt}} , \textbf{EU vaccine strategy} 

\bibitem{reuters}\textit{\href{https://graphics.reuters.com/world-coronavirus-tracker-and-maps/vaccination-rollout-and-access/}{https://graphics.reuters.com/world-coronavirus-tracker-and-maps/vaccination-rollout-and-access/}}, \textbf{Covid-19 vaccination tracker, Reuters} 


\bibitem{pubINE}\textit{\href{https://www.ine.pt/xurl/pub/71882686}{https://www.ine.pt/xurl/pub/71882686}}, \textbf{Instituto Nacional de Estatística - Estatísticas Demográficas : 2019. Lisboa : INE, 2020.} 


\bibitem{trab}\textit{\href{https://ec.europa.eu/eures/main.jsp?acro=lmi&lang=pt&countryId=PT&catId=57&parentId=0}{https://ec.europa.eu/eures/main.jsp?acro=lmi&lang=pt&countryId=PT&catId=57&parentId=0}}, \textbf{Labor Market Information - European Commission} 


\bibitem{dataDownload}\textit{\href{https://vaccinetracker.ecdc.europa.eu/public/extensions/COVID-19/vaccine-tracker.html#notes-tab}} , \textbf{General Vaccine information - European Centre for Disease Prevention and Control } 

\end{thebibliography}

\end{document}


